\section{Analysis}
\label{sec:analysis}

Unlike programming desktops, where one mainly 
  improves software by increasing either features or performance,
  mobile programmers develop for increase features, performance, and 
  battery life.
In fact, one of the main selling points for hetrogenous
  programming on mobile devices, is the increase in battery life.
Increasingly, hardware vendors, such as Apple or Samsung, sell new
  mobile hardware by advertising longer battery life.
Finding a balance between energy consumption and performance is a 
  balancing act that a programmer would like to deligate to the 
  compiler or runtime.


\begin{table}
\centering
\begin{tabular}{ | l | p{2.1cm} | p{1.7cm} | c |}
    \hline 
    Name & CPU & GPU & Memory \\ \hline
    GalaxyNexus & ARMv7, 2 cores, 1200 Mhz, SIMD NEON & PowerVR-SGX 540 & 694Mb \\ \hline
    Nexus5 & Qualcomm Snapdragon S4 Pro 1.5GHz & Adreno 320 400MHz & 2Gb \\ \hline
    Nexus7 & Qualcomm Snapdragon 800 2.26GHz & Adreno 330 450MHz & 2Gb \\ \hline
    Nexus10 & DualCore 1.7GHz Cortex-A15 & Mali T604 & 2Gb \\ \hline
    SM-T900 & QuadCore 1.9GHz Cortex-A15 & Mali T628 & 3Gb \\ \hline
    \hline
\end{tabular}
\caption{The device hardware specifications used for the analysis.}
\label{table:hardware}
\end{table}

In this section evaluate each benchmark on $5$ devices shown
  in Table~\ref{table:hardware} all running Android $4.4.2$ (KitKat).
These devices capture the low, mid, and high end mobile and tablet
  devices that are currently available on the market.
Only the Nexus 5 and Nexus 7 devices can be modified to allow one to install an
  unofficial OpenCL implementation and only those two devices are fully supported
  by Trepn.
Each benchmarks contains multiple implementation which 
  is analyzed to determine tradeoff between
  processor utlization, performance, and energy usage.
We conclude the section by discussing the programability of RenderScript
  and compare it against established programming models such as:
  native C, OpenMP, and OpenCL.

\input{load}
\input{performance}
\input{power}
\input{io}

\input{programability}